---
title: 'Stats 250: Lab 1 Assignment'
author: "Replace this text with your name in quotes"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Instructions

Complete the lab exercises and Dive Deeper questions below. Lab exercises will be graded out of 2 points each and the Dive Deeper questions will be graded out of 3 points each - all of them will be graded for *correctness*. Upon completion, follow the submission instructions at the bottom of this document. We strongly encourage you to work on this assignment with a partner, but each student must submit their own lab assignment.

***

### Background

This data set contains information on *all* 801 Pokemon from the first Seven Generations of Pokemon. The variables in the data set include:

1. `primary_type`: primary type of the Pokemon (grass, fire, water, etc.)
2. `secondary_type`: secondary type of the Pokemon (poison, flying, none, etc.)
3. `legendary_status`: denotes if the Pokemon is of legendary status
4. `attack`: base attack of the Pokemon (in points)
5. `defense`: base defense of the Pokemon (in points)
6. `hp`: base HP (hit points) of the Pokemon (in points)
7. `speed`: base speed of the Pokemon (in points)
8. `special_attack`: base special attack of the Pokemon (in points)
9. `special_defense`: base special defense of the Pokemon (in points)
10. `capture_rate`: capture rate of the Pokemon (in points)
11. `base_total`: total base score of the Pokemon (in points)
12. `generation`: classifies the generation in which the Pokemon was introduced

[Source](https://www.kaggle.com/datasets/rounakbanik/pokemon)

Use the following code chunk to read in the data set. 

```{r readpokemon}
pokemon <- read.csv("pokemon.csv", row.names = 1)
```

And be sure to load in the required packages!

```{r loadPackages}
library(ggplot2)
```


***


### Lab Exercises

After reading the background above, complete the following lab exercises. Remember, these exercises are worth 2 points each and are graded for *correctness*.

> **1.** Select a *quantitative* variable from the data set and create a histogram. Be sure to appropriately update the plot title and axis labels. We highly recommend that you return to the lab01_resources document for example code and help! 

```{r exercise1}
# Replace this comment with your code

```



> **2.** Using the same variable from Exercise 1, calculate at least three useful numerical summaries that would help provide a thorough description of the variable's distribution (to be done in Exercise 3).

```{r exercise2}
# Replace this comment with your code

```



> **3.** Using the histogram from Exercise 1 and the *exact* numerical values calculated in Exercise 2, thoroughly describe the distribution of your quantitative variable. Your response should be 2-3 sentences in length and describe at minimum the distribution's shape, center, and spread. 

*Answer:* Replace this text with your answer.



> **4.** Select a *different* quantitative variable from the data set and create a histogram. Be sure to appropriately update the plot title and axis labels. 

```{r exercise4}
# Replace this comment with your code

```



> **5.** Using the same variable from Exercise 4, calculate at least three useful numerical summaries that would help provide a thorough description of the variable's distribution (to be done in Exercise 6).

```{r exercise5}
# Replace this comment with your code

```



> **6.** Using the histogram from Exercise 4 and the exact numerical values calculated in Exercise 5, thoroughly describe the distribution of your quantitative variable. Your response should be 2-3 sentences in length and describe at minimum the distribution's shape, center, and spread. 

*Answer:* Replace this text with your answer.



> **7.** Select a *categorical* variable from the data set and create a bar chart in the first code chunk. Be sure to appropriately update the plot title and axis labels. In the second code chunk, create a one-way frequency table for your categorical variable.

```{r exercise7_barChart}
# Replace this comment with your code

```

```{r exercise7_table}
# Replace this comment with your code

```



***


### Dive Deeper Questions

After completing the lab exercises above, complete the following Dive Deeper questions. Remember, these questions are worth 3 points each and are graded for correctness.

> **1.** The last variable in the data set is `generation`. This variable classifies the generation in which the Pokemon was introduced. Would this variable be considered quantitative or categorical? Explain your answer in 2-3 sentences.

[Wikipedia](https://en.wikipedia.org/wiki/Pok%C3%A9mon_(video_game_series))

*Answer:* Replace this text with your answer.



> **2.** Propose a question that could be answered using the Pokemon data set. Your question should involve *two* variables. After stating your question, identify which variable is acting as the response variable and which variable is acting as the explanatory variable.

*Answer:* Replace this text with your answer.



***



### Submission Instructions

Once you have finished, follow the instructions below to submit your work.

1. At the top of the document, change the `author` field to your name (in quotes!). **Do not change any of the other fields (title, date, output) at the top of the document.**

2. Click the Knit button to create an HTML file of your work. If you get an error that says "Duplicate chunk label", go back and check your code chunk names - each one must have a unique name!  

3. In the Files pane (bottom right window), check the box next to `lab01_assignment.html`. 

4. In the Files pane, click on the gear icon (More) and select "Export..."  

5. Do not change the file name (`lab01_assignment.html`). Click Download and save the file to your computer. 

6. On the Stats 250 Canvas site, navigate to the Lab 1 assignment. 

7. Upload your saved `lab01_assignment.html` file. 

8. Click "Submit Assignment". Congratulations, you are done with Lab 1!


